Timing and Positioning: Insights from Steak-Related Product Pricing at Walmart and Loblaws

### 2024-12-03 paper/paper.qmd
Update paper.qmd

```{r}
#| label: box-plot
#| fig-cap: Current prices of steak-related products by vendor and month, showing Walmart's consistent lower pricing and Loblaws' broader range with seasonal price fluctuations.
#| echo: false
#### Read data ####
analysis_data <- read_parquet("/Users/lucky/Desktop/Explore-steak-related-products/data/02-analysis_data/analysis_data.parquet")

@@

#### Distribution of current_price by vendor ####
ggplot(analysis_data, aes(x = reorder(vendor, current_price, FUN = median), y = current_price, fill = vendor)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  coord_flip() + # Flip the coordinates for better readability
  labs(
    title = "Distribution of Current Price by Vendor",
    subtitle = "Steak-related products",
    x = "Vendor",
    y = "Current Price (CAD)",
    fill = "Vendor"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

#### Distribution of current_price by month ####
ggplot(analysis_data, aes(x = month, y = current_price, fill = month)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(
    title = "Monthly Distribution of Current Price",
    subtitle = "Steak-related products",
    x = "Month",
    y = "Current Price (CAD)",
    fill = "Month"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

Together, these visualizations reveal insights into vendor-specific pricing strategies and seasonal variations, critical for understanding consumer and market behavior in the steak-related product segment.
```{r}
#| label: scatter-plot
#| fig-cap: Scatter plot showing a positive relationship between old and current prices, with outliers indicating significant markdowns on premium products.
#| echo: false
#| warning: false
#| message: false

@@

#### Distribution of current_price by old_price ####
ggplot(analysis_data, aes(x = old_price, y = current_price)) +
  geom_point(alpha = 0.6, color = "orange", size = 2) + # Add transparency and size for clarity
  geom_smooth(method = "lm", color = "purple", se = FALSE, linetype = "dashed") + # Add trendline
  labs(
    title = "Relationship Between Old and Current Prices",
    subtitle = "Steak-related products",
    x = "Old Price (CAD)",
    y = "Current Price (CAD)"
  ) +
  theme_light() + # Change to a cleaner theme
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

>>>>

```{r}
#| label: fig-bll1
#| fig-cap: Current prices of steak-related products by vendor and month, showing Walmart's consistent lower pricing and Loblaws' broader range with seasonal price fluctuations.
#| echo: false
#### Read data ####
analysis_data <- read_parquet("/Users/lucky/Desktop/Explore-steak-related-products/data/02-analysis_data/analysis_data.parquet")

@@

#### Distribution of current_price by vendor with actual data overlay ####
ggplot(analysis_data, aes(x = reorder(vendor, current_price, FUN = median), y = current_price, fill = vendor)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, alpha = 0.7) +  # Add transparency to boxplots
  geom_jitter(aes(color = vendor), width = 0.2, size = 1.5, alpha = 0.5) +  # Overlay actual data points with jitter for clarity
  coord_flip() +  # Flip the coordinates for better readability
  labs(
    title = "Distribution of Current Price by Vendor",
    subtitle = "Steak-related products",
    x = "Vendor",
    y = "Current Price (CAD)",
    fill = "Vendor",
    color = "Actual Data"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

#### Distribution of current_price by month with actual data overlay ####
ggplot(analysis_data, aes(x = month, y = current_price, fill = month)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2, alpha = 0.7) +  # Add transparency to boxplots
  geom_jitter(aes(color = month), width = 0.2, size = 1.5, alpha = 0.5) +  # Overlay actual data points with jitter for clarity
  labs(
    title = "Monthly Distribution of Current Price",
    subtitle = "Steak-related products",
    x = "Month",
    y = "Current Price (CAD)",
    fill = "Month",
    color = "Actual Data"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

Together, these visualizations reveal insights into vendor-specific pricing strategies and seasonal variations, critical for understanding consumer and market behavior in the steak-related product segment.
```{r}
#| label: fig-bll2
#| fig-cap: Scatter plot showing a positive relationship between old and current prices, with outliers indicating significant markdowns on premium products.
#| echo: false
#| warning: false
#| message: false

@@

#### Distribution of current_price by old_price ####
ggplot(analysis_data, aes(x = old_price, y = current_price)) +
  geom_point(alpha = 0.4, color = "orange", size = 3) + # Adjusted transparency and point size for better visibility
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") + # Changed trendline to solid for better contrast
  labs(
    title = "Relationship Between Old and Current Prices",
    subtitle = "Steak-related Products",
    x = "Old Price (CAD)",
    y = "Current Price (CAD)"
  ) +
  theme_minimal() + # Use a minimal theme for better focus on data
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5), # Center-align title
    plot.subtitle = element_text(size = 12, hjust = 0.5), # Center-align subtitle
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )
